Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-12-07 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
change group name

%\begin{equation}
%\sum  S_{p}-1
%\label{saturation_constrain}
%\end{equation}
\noindent
where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $n$ is the current time level. $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. In addition to the numerical scheme above the discretised global mass and force balance equations are solved using a multigrid-like approach \citep[see][]{pavlidis2016}. This numerical formulation is fully described by \citet{gomes_2017}\citep[see also][]{salinas2015,adam_2016}.

@@

\medskip 
All test-cases were performed in idealised geometries discretised with unstructured triangular mesh using the \PN[1]{2} and \PN[1]{1} FE-pairs. An implicit Crank-Nicolson time-stepping scheme was used with an {\it a posteriori} adaptive time-step size targeting a maximum Courant-Friedrichs-Lewy \citep{courant_1941} number of 2. In most simulations, the domain was initially fully saturated with non-wetting fluid which was displaced by a pure (wetting) fluid at a prescribed initial velocity $\left(u^{0}\right)$. For simplicity, the porosity ($\phi$) of the domain was kept constant at 0.2 in all simulations, whereas the absolute permeability ($\mathbf{K}$) varied in space, \ie $\mathbf{K}=\mathbf{K}\left(x_{i}\right)$. Fluids are assumed incompressible and gravity was neglected. The set-up for the numerical simulations are summarised in Table \ref{table:setup}.

>>>>

%\begin{equation}
%\sum  S_{p}-1
%\label{saturation_constrain}
%\end{equation}
\noindent
where $\Omega_{CVi}$ and $\Gamma_{CVi}$ are the volume and boundary of CV i respectively, \textbf{n} is the outward pointing unit normal vector to the surface of $CV_{i}$ and $n$ is the current time level. $\theta$ varies smoothly between $0.5$ (corresponding to Crank-Nicolson method) and $1$ (corresponding to backward-Euler scheme) to avoid the introduction of spurious oscillations for large grid Courant numbers. The discretised global mass and force balance equations (Eqns.~\ref{force-semi-disc}-\ref{global_mass_balance}) are solved using a multigrid-like approach as described by \citet{pavlidis2016} \citep[see also][]{salinas_2016}. The numerical formulation is fully described by \citet{gomes_2017} \citep[see also][]{salinas2015,adam_2016}.

@@

\medskip 
All test-cases were performed in idealised geometries discretised with unstructured triangular mesh using the \PN[1]{2} and \PN[1]{1} FE-pairs (Fig.~\ref{fig:fem_elem}). An implicit Crank-Nicolson time-stepping scheme was used with an {\it a posteriori} adaptive time-step size targeting a maximum Courant-Friedrichs-Lewy \citep{courant_1941} number of 2. In most simulations, the domain was initially fully saturated with non-wetting fluid which was displaced by a pure (wetting) fluid at a prescribed initial velocity $\left(u^{0}\right)$. For simplicity, the porosity ($\phi$) of the domain was kept constant at 0.2 in all simulations, whereas the absolute permeability ($\mathbf{K}$) varied in space, \ie $\mathbf{K}=\mathbf{K}\left(x_{i}\right)$. Fluids are assumed incompressible and gravity was neglected. The set-up for the numerical simulations are summarised in Table \ref{table:setup}.
